---
title: "Week 4 assignment"
author: "Jonathan Roberts"
date: "31 March 2017"
output: 
    html_notebook: default
    html_document: default
---

#Get up and running
##Global RMarkdown Parameters
Are set here

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

##Install the R markdown packages and library

```{r message = FALSE, warning = TRUE}
libraries <- c("rmarkdown")

#UNCOMMENT TO INSTALL PACKAGES
#lapply(libraries, install.packages)
lapply(libraries, library, character.only = TRUE)
```


## Working Directory: 
Set up where you are working and what is in the directory and what objects you have
 
```{r}

# SET THE WORKING DIRECTORY
setwd("C:/DataScience/John_Hopkins")
dir()
ls()

```

#Assignment: Part 1

```{r}

best <- function(state,outcome){
  
  ##Read in the data
  outcomes <- read.csv("hospital_data/outcome-of-care-measures.csv", colClasses = "character")

  ##Set the columns where the data is - and the master list of conditions
  conditions<-data.frame( c("heart attack","heart failure","pneumonia"),c(11,17,23))

  #find out if the call is valid
  state_test<-is.na(match(state,outcomes$State)) ## returns TRUE for an ERROR
  condition_test<- is.na(match(outcome,conditions[,1])) ## returns TRUE for an ERROR
  valid_call<-condition_test | state_test ## returns TRUE for an ERROR

  ##TODO - WRITE THE STOP LINE TO EXIT THE FUNCTION
  
  #Make my data numeric
  
  #data<-conditions[,2]
  ####outcomes[,data]<-as.numeric(outcomes[,data]) ##TODO - why doesn't this work??
  outcomes[,11]<-as.numeric(outcomes[,11])
  outcomes[,17]<-as.numeric(outcomes[,17])  
  outcomes[,23]<-as.numeric(outcomes[,23])
  
  # SPLIT BY STATE: THIS IS QUITE RESOURCE INTENSIVE - MIGHT BE BEST NOT TO 'FILTER' ON STATE NAME.
  outcome_by_state<-split(outcomes, outcomes$State)
  #PUL OUT THE SELECTED STATE DAATA AS A DATA FRAME
  state_df<-outcome_by_state[[state]]
  
  #READY TO 'SORT' - ORDER SEEMS TO BE THE BEST...
  cond<-conditions[match(outcome,conditions[,1]),2]  
  order_by_MR<-state_df[order(state_df[,cond],state_df[,2]),]
  order_by_MR[1,2]
  
}

  best("TX","heart failure")

```



Write in the data

```{r}

outcome <- read.csv("hospital_data/outcome-of-care-measures.csv", colClasses = "character")
head(outcome)
str(outcome)
summary(outcome)
class(outcome)
colnames(outcome)
outcome[,11]
ncol(outcome)
nrow(outcome)
outcome[, 11] <- as.numeric(outcome[, 11])
## You may get a warning about NAs being introduced; that is okay
hist(outcome[, 11])
head(outcome)

```

## Do the main bit of ordering and finding number 1
```{r}
#Where is the data?
heart_attack<-outcome[,11]

outcome[,17]<-as.numeric(outcome[,17])
heart_failure<-outcome[,17]
hist(outcome[, 17])

outcome[,23]<-as.numeric(outcome[,23])
pneumonia<-outcome[,23]
hist(outcome[, 23])


# SPLIT BY STATE: THIS IS QUITE RESOURCE INTENSIVE - MIGHT BE BEST NOT TO SPLIT AND 'FILTER' ON STATE NAME.
outcome_by_state<-split(outcome, outcome$State)
class(outcome_by_state)
str(outcome_by_state)
#PUL OUT THE SELECTED STATE DAATA AS A DATA FRAME
class(outcome_by_state$AK)
ak_df<-outcome_by_state$AK
ak_df

#READY TO 'SORT' - ORDER SEEMS TO BE THE BEST...

order_by_MR<-ak_df[order(ak_df[,11],ak_df[,2]),]
order_by_MR[1,]

```

##Do the error trapping
```{r}
#find out if we are one of the states

state<-"AK"
condition<-"heart attack"
conditions<-data.frame( c("heart attack","heart failure","pneumonia"),c(11,17,23))

state_test<-is.na(match(state,outcome$State)) ## returns TRUE for an ERROR
condition_test<- is.na(match(condition,conditions[,1])) ## returns TRUE for an ERROR

valid_call<-condition_test | state_test ## returns TRUE for an ERROR


```




```{r}

pollutantmean <- function(directory, pollutant, id = 1:332){
    ##directory is a character vector of length 1 indicating the location of the CCSV        files
    ## pollutant is a character vector of length 1 indicating which pollutant we want to call data on: either sulfate or nitrate
    ## id is the intedger vector indicating which monitor station we want
    ## overall function returns the mean of the pollutant, across all files - ignoring NA's
   
 
    #Fuck about making the adding 00's to numbers <100 then paste with id
    len_00=length(id[id<10])
    len_0=length(id[id>=10&id<100])
    len=length(id)-len_0-len_00
    zeros<-c(rep("00", times = len_00), rep("0", times = len_0), rep("", times = len))
   
    # A character vector with the 'filename'
    full_id<-paste(zeros,id,sep = "")
    

    #Now make the full filename in a vector
    getwd()    
    folder<-paste(getwd(),directory,full_id, sep = "/")
    filename<-paste(folder,"csv", sep=".")
    
    # Now loop around the files reading out into 1 big dataframe
    # i've used the function rbind.data.frame which seems to be exactly what i need.
    ##TODO, not sure i really need the if (i==1) loop or i can 'initialise' another way.

    for (i in seq_along(id)) {
        pollute_data<-read.csv(file = filename[i])
        if(i == 1){
            full_data_df<-pollute_data
            check_row<-nrow(pollute_data)
        }
        else{
            full_data_df<-rbind.data.frame(full_data_df,pollute_data)
            check_row <- check_row + nrow(pollute_data)    
        }
    }
    #TEST IF WE HAVE EQUAL rows i nour data and our data.frame
    nrow(full_data_df) == check_row
    full_data_choice<-full_data_df[[pollutant]]
    mean(full_data_choice, na.rm = TRUE)
    
}

pollutantmean("specdata", "nitrate", id = 23)
 

```


#Part 2

Write a function that reads a directory full of files and reports the number of completely observed cases in each data file. The function should return a data frame where the first column is the name of the file and the second column is the number of complete cases. A prototype of this function follows


You can see some example output from this function. The function that you write should be able to match this output. Please save your code to a file named complete.R. To run the submit script for this part, make sure your working directory has the file complete.R in it.

```{r}
complete <- function(directory, id = 1:332){
    ## directory is a character vector of length 1 indicating the location of the CSV files
    ## id is the intedger vector indicating which monitor station we want
  
    #Fuck about making the adding 00's to numbers <100 then paste with id
    len_00=length(id[id<10])
    len_0=length(id[id>=10&id<100])
    len=length(id)-len_0-len_00
    
    zeros<-c(rep("00", times = len_00), rep("0", times = len_0), rep("", times = len))
    full_id<-paste(zeros,id,sep = "")
    

    #first data set
    getwd()    
    folder<-paste(getwd(),directory,full_id, sep = "/")
    filename<-paste(folder,"csv", sep=".")
    
    # filename # this is a vector of filenames!
    
    ## Above here is vectorised, below is not#########

    ##TODO, not sure i really need the if (i==1) loop or if I can 'initialise' another way.

    for (i in seq_along(id)) {
        pollute_data<-read.csv(file = filename[i])
        both_df<-pollute_data[(!is.na(pollute_data$sulfate) & !is.na(pollute_data$nitrate))==TRUE,]
        if(i == 1){
            full_data_df<-pollute_data
            check_row<-nrow(pollute_data)
            complete_summary <- matrix(c(pollute_data[1,"ID"],nrow(both_df)), nrow=1,ncol=2)   
            colnames(complete_summary) <- c("id","nobs")
        }
        else{
            full_data_df<-rbind.data.frame(full_data_df,pollute_data)
            check_row <- check_row + nrow(pollute_data)
            complete_summary<- rbind(complete_summary, c(pollute_data[1,"ID"],nrow(both_df)))
        }
    }
    #TEST IF WE HAVE EQUAL rows i nour data and our data.frame
    nrow(full_data_df) == check_row
    
    # output the summary of compelte data
    class(complete_summary)    
    complete_summary
    
}

complete("specdata",30:25)

```


#Part 3

Write a function that takes a directory of data files and a threshold for complete cases and calculates the correlation between sulfate and nitrate for monitor locations where the number of completely observed cases (on all variables) is greater than the threshold. The function should return a vector of correlations for the monitors that meet the threshold requirement. If no monitors meet the threshold requirement, then the function should return a numeric vector of length 0. A prototype of this function follows

For this function you will need to use the 'cor' function in R which calculates the correlation between two vectors. Please read the help page for this function via '?cor' and make sure that you know how to use it.

You can see some example output from this function. The function that you write should be able to match this output. Please save your code to a file named corr.R. To run the submit script for this part, make sure your working directory has the file corr.R in it.



```{r}

corr <- function(directory, threshold = 0){
    ## directory is a character vector of length 1 indicating the location of the CSV files
    ## threshold is the numeric vector indicating the minimum number of completely observed observations to compute the correlations.
    
    
    #directory<-"specdata"
    id<-1:332
    #threshold =  150 
    
    #Fuck about making the adding 00's to numbers <100 then paste with id
    len_00=length(id[id<10])
    len_0=length(id[id>=10&id<100])
    len=length(id)-len_0-len_00
    
    zeros<-c(rep("00", times = len_00), rep("0", times = len_0), rep("", times = len))
    full_id<-paste(zeros,id,sep = "")

    #first data set
    folder<-paste(getwd(),directory,full_id, sep = "/")
    filename<-paste(folder,"csv", sep=".")

    #Loop around reading in data; then store it and do a bit of analysis
    ##Could this be done without the If i==1 loop? 
    for (i in seq_along(id)) {
        pollute_data<-read.csv(file = filename[i])
        both_df<-pollute_data[(!is.na(pollute_data$sulfate) & !is.na(pollute_data$nitrate))==TRUE,]
        corr_vec<-cor(both_df$sulfate, both_df$nitrate)
        if(i == 1){
            #full_data_df<-pollute_data
            #check_row<-nrow(pollute_data)
            complete_data_df<-both_df
            complete_summary <- matrix(c(pollute_data[1,"ID"],nrow(both_df),corr_vec), nrow=1,ncol=3) 
            colnames(complete_summary) <- c("id","nobs", "corr")
        }
        else{
            #full_data_df<-rbind.data.frame(full_data_df,pollute_data)
            #check_row <- check_row + nrow(pollute_data)
            complete_data_df<-rbind.data.frame(complete_data_df,both_df)
            complete_summary<- rbind(complete_summary, c(pollute_data[1,"ID"],nrow(both_df),corr_vec))
        }
    }
    
    test<-complete_summary[,2]>threshold
    #complete_summary[test] ## TODO: Why the hell does this lose it's matrixness??????
    threshold_summary<-matrix(complete_summary[test],ncol=3)
    threshold_summary[,3]
    
}
    

cr<-corr("specdata",0)
head(cr)
summary(cr)

```




# Sandbox 
Place to experiment before dropping something that works into the function.

## Part 1 sandbox

```{r}
 directory<-"specdata"
    id<-70:72
    pollutant<-"nitrate"
    
    #Fuck about making the adding 00's to numbers <100 then paste with id
    len_00=length(id[id<10])
    len_0=length(id[id>=10&id<100])
    len=length(id)-len_0-len_00
    
    zeros<-c(rep("00", times = len_00), rep("0", times = len_0), rep("", times = len))
    full_id<-paste(zeros,id,sep = "")
    

    #first data set
    getwd()    
    folder<-paste(getwd(),directory,full_id, sep = "/")
    filename<-paste(folder,"csv", sep=".")
    
    # filename # this is a vector of filenames!
    
    ## Above here is vectorised, below is not#########

    ##Function still doesn;t really work for numbers <100
    ##TODO, not sure i really need the if (i==1) loop or i can 'initialise' another way.

    for (i in seq_along(id)) {
        pollute_data<-read.csv(file = filename[i])
        if(i == 1){
            full_data_df<-pollute_data
            check_row<-nrow(pollute_data)
        }
        else{
            full_data_df<-rbind.data.frame(full_data_df,pollute_data)
            check_row <- check_row + nrow(pollute_data)    
        }
    }
    #TEST IF WE HAVE EQUAL rows i nour data and our data.frame
    nrow(full_data_df) == check_row
    full_data_choice<-full_data_df[[pollutant]]
    mean(full_data_choice, na.rm = TRUE)
    
            
```


## Part 2 sandbox
Write a function that reads a directory full of files and reports the number of completely observed cases in each data file. The function should return a data frame where the first column is the name of the file and the second column is the number of complete cases. A prototype of this function follows

complete <- function(directory, id = 1:332)
    ## directory is a character vector of length 1 indicating the location of the CSV files
    ## id is the intedger vector indicating which monitor station we want

    ## General logic:
    # Read in one file
    # work out all the obeserved files probably using any or all
    # count for that file
    # loop around
   


```{r}
    directory<-"specdata"
    id<-30:25
    
    #Fuck about making the adding 00's to numbers <100 then paste with id
    len_00=length(id[id<10])
    len_0=length(id[id>=10&id<100])
    len=length(id)-len_0-len_00
    
    zeros<-c(rep("00", times = len_00), rep("0", times = len_0), rep("", times = len))
    full_id<-paste(zeros,id,sep = "")
    

    #first data set
    getwd()    
    folder<-paste(getwd(),directory,full_id, sep = "/")
    filename<-paste(folder,"csv", sep=".")
    
    # filename # this is a vector of filenames!
    
    ## Above here is vectorised, below is not#########

    ##TODO, not sure i really need the if (i==1) loop or if I can 'initialise' another way.

    for (i in seq_along(id)) {
        pollute_data<-read.csv(file = filename[i])
        both_df<-pollute_data[(!is.na(pollute_data$sulfate) & !is.na(pollute_data$nitrate))==TRUE,]
        if(i == 1){
            full_data_df<-pollute_data
            check_row<-nrow(pollute_data)
            complete_summary <- matrix(c(pollute_data[1,"ID"],nrow(both_df)), nrow=1,ncol=2)   
            colnames(complete_summary) <- c("id","nobs")
        }
        else{
            full_data_df<-rbind.data.frame(full_data_df,pollute_data)
            check_row <- check_row + nrow(pollute_data)
            complete_summary<- rbind(complete_summary, c(pollute_data[1,"ID"],nrow(both_df)))
        }
    }
    #TEST IF WE HAVE EQUAL rows i nour data and our data.frame
    nrow(full_data_df) == check_row
    
    # output the summary of compelte data
    class(complete_summary)    
    complete_summary
            
```
```{r}
   # in a couple of lines
    both <- !is.na(full_data_df$sulfate) & !is.na(full_data_df$sulfate)
    full_data_df[315,]
    full_data_df[both==TRUE,]
    length(both[both==TRUE])
    
    #in one line
    both_df <- full_data_df[(!is.na(full_data_df$sulfate) & !is.na(full_data_df$sulfate))==TRUE,]
    nrow(both_df)
```


#Part 3 Sandbox

Write a function that takes a directory of data files and a threshold for complete cases and calculates the correlation between sulfate and nitrate for monitor locations where the number of completely observed cases (on all variables) is greater than the threshold. The function should return a vector of correlations for the monitors that meet the threshold requirement. If no monitors meet the threshold requirement, then the function should return a numeric vector of length 0. A prototype of this function follows

For this function you will need to use the 'cor' function in R which calculates the correlation between two vectors. Please read the help page for this function via '?cor' and make sure that you know how to use it.


##Testing cor
```{r}

x<-matrix(1:20, nrow = 10, ncol = 2)
x
cor(x[,1], x[,2])

```

##Writing cor with no threshold

```{r}
corr <- function(directory, threshold = 0, id = 1:332){
    ## directory is a character vector of length 1 indicating the location of the CSV files
    ## threshold is the numeric vector indicating the minimum number of completely observed observations to compute the correlations.
    
    
    # In the cor function you can pass both variable together in a matrix/data frame
    
}
    
```


```{r}
    directory<-"specdata"
    id<-1:332
    threshold =  150 ## Currently unused
    
    #Fuck about making the adding 00's to numbers <100 then paste with id
    len_00=length(id[id<10])
    len_0=length(id[id>=10&id<100])
    len=length(id)-len_0-len_00
    
    zeros<-c(rep("00", times = len_00), rep("0", times = len_0), rep("", times = len))
    full_id<-paste(zeros,id,sep = "")

    #first data set
    folder<-paste(getwd(),directory,full_id, sep = "/")
    filename<-paste(folder,"csv", sep=".")

    for (i in seq_along(id)) {
        pollute_data<-read.csv(file = filename[i])
        both_df<-pollute_data[(!is.na(pollute_data$sulfate) & !is.na(pollute_data$nitrate))==TRUE,]
        corr_vec<-cor(both_df$sulfate, both_df$nitrate)
        if(i == 1){
            #full_data_df<-pollute_data
            #check_row<-nrow(pollute_data)
            complete_data_df<-both_df
            complete_summary <- matrix(c(pollute_data[1,"ID"],nrow(both_df),corr_vec), nrow=1,ncol=3) 
            colnames(complete_summary) <- c("id","nobs", "corr")
        }
        else{
            #full_data_df<-rbind.data.frame(full_data_df,pollute_data)
            #check_row <- check_row + nrow(pollute_data)
            complete_data_df<-rbind.data.frame(complete_data_df,both_df)
            complete_summary<- rbind(complete_summary, c(pollute_data[1,"ID"],nrow(both_df),corr_vec))
        }
    }
    
    test<-complete_summary[,2]>threshold
    #complete_summary[test] ## TODO: Why the hell does this lose it's matrixness??????
    threshold_summary<-matrix(complete_summary[test],ncol=3)
    threshold_summary[,3]

    
    
            
```




