---
title: "Week 2 assignment"
author: "Jonathan Roberts"
date: "24 March 2017"
output: 
    html_notebook: default
    html_document: default
---

#Get up and running
##Global RMarkdown Parameters
Are set here

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

##Install the R markdown packages and library

```{r message = FALSE, warning = TRUE}
libraries <- c("rmarkdown")

#UNCOMMENT TO INSTALL PACKAGES
#lapply(libraries, install.packages)
lapply(libraries, library, character.only = TRUE)
```


## Working Directory: 
Set up where you are working and what is in the directory and what objects you have

```{r}

# SET THE WORKING DIRECTORY
setwd("/Users/jonathanroberts/Documents/R/John_Hopkins_Rbasics")
dir()
ls()

```

#Assignment: Part 1

Write a function named 'pollutantmean' that calculates the mean of a pollutant (sulfate or nitrate) across a specified list of monitors. The function 'pollutantmean' takes three arguments: 'directory', 'pollutant', and 'id'. Given a vector monitor ID numbers, 'pollutantmean' reads that monitors' particulate matter data from the directory specified in the 'directory' argument and returns the mean of the pollutant across all of the monitors, ignoring any missing values coded as NA. A prototype of the function is as follows

```{r}

pollutantmean <- function(directory, pollutant, id = 1:332){
    ##directory is a character vector of length 1 indicating the location of the CCSV        files
    ## pollutant is a character vector of length 1 indicating which pollutant we want to call data on: either sulfate or nitrate
    ## id is the intedger vector indicating which monitor station we want
    ## overall function returns the mean of the pollutant, across all files - ignoring NA's
   
 
    #Fuck about making the adding 00's to numbers <100 then paste with id
    len_00=length(id[id<10])
    len_0=length(id[id>=10&id<100])
    len=length(id)-len_0-len_00
    zeros<-c(rep("00", times = len_00), rep("0", times = len_0), rep("", times = len))
   
    # A character vector with the 'filename'
    full_id<-paste(zeros,id,sep = "")
    

    #Now make the full filename in a vector
    getwd()    
    folder<-paste(getwd(),directory,full_id, sep = "/")
    filename<-paste(folder,"csv", sep=".")
    
    # Now loop around the files reading out into 1 big dataframe
    # i've used the function rbind.data.frame which seems to be exactly what i need.
    ##TODO, not sure i really need the if (i==1) loop or i can 'initialise' another way.

    for (i in seq_along(id)) {
        pollute_data<-read.csv(file = filename[i])
        if(i == 1){
            full_data_df<-pollute_data
            check_row<-nrow(pollute_data)
        }
        else{
            full_data_df<-rbind.data.frame(full_data_df,pollute_data)
            check_row <- check_row + nrow(pollute_data)    
        }
    }
    #TEST IF WE HAVE EQUAL rows i nour data and our data.frame
    nrow(full_data_df) == check_row
    full_data_choice<-full_data_df[[pollutant]]
    mean(full_data_choice, na.rm = TRUE)
    
}

pollutantmean("specdata", "nitrate", id = 23)
 

```


#Part 2

Write a function that reads a directory full of files and reports the number of completely observed cases in each data file. The function should return a data frame where the first column is the name of the file and the second column is the number of complete cases. A prototype of this function follows


You can see some example output from this function. The function that you write should be able to match this output. Please save your code to a file named complete.R. To run the submit script for this part, make sure your working directory has the file complete.R in it.

```{r}
complete <- function(directory, id = 1:332){
    ## directory is a character vector of length 1 indicating the location of the CSV files
    ## id is the intedger vector indicating which monitor station we want
  
    #Fuck about making the adding 00's to numbers <100 then paste with id
    len_00=length(id[id<10])
    len_0=length(id[id>=10&id<100])
    len=length(id)-len_0-len_00
    
    zeros<-c(rep("00", times = len_00), rep("0", times = len_0), rep("", times = len))
    full_id<-paste(zeros,id,sep = "")
    

    #first data set
    getwd()    
    folder<-paste(getwd(),directory,full_id, sep = "/")
    filename<-paste(folder,"csv", sep=".")
    
    # filename # this is a vector of filenames!
    
    ## Above here is vectorised, below is not#########

    ##TODO, not sure i really need the if (i==1) loop or if I can 'initialise' another way.

    for (i in seq_along(id)) {
        pollute_data<-read.csv(file = filename[i])
        both_df<-pollute_data[(!is.na(pollute_data$sulfate) & !is.na(pollute_data$nitrate))==TRUE,]
        if(i == 1){
            full_data_df<-pollute_data
            check_row<-nrow(pollute_data)
            complete_summary <- matrix(c(pollute_data[1,"ID"],nrow(both_df)), nrow=1,ncol=2)   
            colnames(complete_summary) <- c("id","nobs")
        }
        else{
            full_data_df<-rbind.data.frame(full_data_df,pollute_data)
            check_row <- check_row + nrow(pollute_data)
            complete_summary<- rbind(complete_summary, c(pollute_data[1,"ID"],nrow(both_df)))
        }
    }
    #TEST IF WE HAVE EQUAL rows i nour data and our data.frame
    nrow(full_data_df) == check_row
    
    # output the summary of compelte data
    class(complete_summary)    
    complete_summary
    
}

complete("specdata",30:25)

```



Part 3

Write a function that takes a directory of data files and a threshold for complete cases and calculates the correlation between sulfate and nitrate for monitor locations where the number of completely observed cases (on all variables) is greater than the threshold. The function should return a vector of correlations for the monitors that meet the threshold requirement. If no monitors meet the threshold requirement, then the function should return a numeric vector of length 0. A prototype of this function follows

For this function you will need to use the 'cor' function in R which calculates the correlation between two vectors. Please read the help page for this function via '?cor' and make sure that you know how to use it.

You can see some example output from this function. The function that you write should be able to match this output. Please save your code to a file named corr.R. To run the submit script for this part, make sure your working directory has the file corr.R in it.



```{r}

corr <- function(directory, threshold = 0){
    ## directory is a character vector of length 1 indicating the location of the CSV files
    ## threshold is the numeric vector indicating the minimum number of completely observed observations to compute the correlations.
    
    
}
    

```



# Sandbox 
Place to experiment before dropping something that works into the function.

## Part 1 sandbox

```{r}
 directory<-"specdata"
    id<-70:72
    pollutant<-"nitrate"
    
    #Fuck about making the adding 00's to numbers <100 then paste with id
    len_00=length(id[id<10])
    len_0=length(id[id>=10&id<100])
    len=length(id)-len_0-len_00
    
    zeros<-c(rep("00", times = len_00), rep("0", times = len_0), rep("", times = len))
    full_id<-paste(zeros,id,sep = "")
    

    #first data set
    getwd()    
    folder<-paste(getwd(),directory,full_id, sep = "/")
    filename<-paste(folder,"csv", sep=".")
    
    # filename # this is a vector of filenames!
    
    ## Above here is vectorised, below is not#########

    ##Function still doesn;t really work for numbers <100
    ##TODO, not sure i really need the if (i==1) loop or i can 'initialise' another way.

    for (i in seq_along(id)) {
        pollute_data<-read.csv(file = filename[i])
        if(i == 1){
            full_data_df<-pollute_data
            check_row<-nrow(pollute_data)
        }
        else{
            full_data_df<-rbind.data.frame(full_data_df,pollute_data)
            check_row <- check_row + nrow(pollute_data)    
        }
    }
    #TEST IF WE HAVE EQUAL rows i nour data and our data.frame
    nrow(full_data_df) == check_row
    full_data_choice<-full_data_df[[pollutant]]
    mean(full_data_choice, na.rm = TRUE)
    
            
```


## Part 2 sandbox
Write a function that reads a directory full of files and reports the number of completely observed cases in each data file. The function should return a data frame where the first column is the name of the file and the second column is the number of complete cases. A prototype of this function follows

complete <- function(directory, id = 1:332)
    ## directory is a character vector of length 1 indicating the location of the CSV files
    ## id is the intedger vector indicating which monitor station we want

    ## General logic:
    # Read in one file
    # work out all the obeserved files probably using any or all
    # count for that file
    # loop around
   


```{r}
    directory<-"specdata"
    id<-30:25
    
    #Fuck about making the adding 00's to numbers <100 then paste with id
    len_00=length(id[id<10])
    len_0=length(id[id>=10&id<100])
    len=length(id)-len_0-len_00
    
    zeros<-c(rep("00", times = len_00), rep("0", times = len_0), rep("", times = len))
    full_id<-paste(zeros,id,sep = "")
    

    #first data set
    getwd()    
    folder<-paste(getwd(),directory,full_id, sep = "/")
    filename<-paste(folder,"csv", sep=".")
    
    # filename # this is a vector of filenames!
    
    ## Above here is vectorised, below is not#########

    ##TODO, not sure i really need the if (i==1) loop or if I can 'initialise' another way.

    for (i in seq_along(id)) {
        pollute_data<-read.csv(file = filename[i])
        both_df<-pollute_data[(!is.na(pollute_data$sulfate) & !is.na(pollute_data$nitrate))==TRUE,]
        if(i == 1){
            full_data_df<-pollute_data
            check_row<-nrow(pollute_data)
            complete_summary <- matrix(c(pollute_data[1,"ID"],nrow(both_df)), nrow=1,ncol=2)   
            colnames(complete_summary) <- c("id","nobs")
        }
        else{
            full_data_df<-rbind.data.frame(full_data_df,pollute_data)
            check_row <- check_row + nrow(pollute_data)
            complete_summary<- rbind(complete_summary, c(pollute_data[1,"ID"],nrow(both_df)))
        }
    }
    #TEST IF WE HAVE EQUAL rows i nour data and our data.frame
    nrow(full_data_df) == check_row
    
    # output the summary of compelte data
    class(complete_summary)    
    complete_summary
            
```
```{r}
   # in a couple of lines
    both <- !is.na(full_data_df$sulfate) & !is.na(full_data_df$sulfate)
    full_data_df[315,]
    full_data_df[both==TRUE,]
    length(both[both==TRUE])
    
    #in one line
    both_df <- full_data_df[(!is.na(full_data_df$sulfate) & !is.na(full_data_df$sulfate))==TRUE,]
    nrow(both_df)
```



